Modelling the Co-evolution of Trade and Culture in Past Societies

### 2015-07-02 article.tex
pr-value

We first analyse the result obtained in the ``neutral model" with one good. The figure~\ref{fig:allMutation} presents the results obtained for two population sizes $N$	 (250 and 500 agents) and $\mu$ varying through three values (0.004,0.016,0.064). The figure is a log-log plot of the average (across all the runs) of the distribution of variants obtained through all experiments. The y-axis shows the frequencies of the variants of the prices used during the simulation, the x-axis shows how many variant achieves such frequencies.

\begin{figure}[!h]
	\centering
	\begin{tabular}{ c c}
		 $N=250$ & $N=500$ \\
		\includegraphics[width=6cm]{img/allmuRandMaxN250.pdf}&
		\includegraphics[width=6cm]{img/allmuRandMaxN500.pdf}
	\end{tabular}
	\caption{Distribution of frequencies depending on the $\mu$ parameter with 250 agents (left) and 500 agents (right). Each plot is the mean obtenaid for 100 runs.\label{fig:allMutation}}
\end{figure}

@@

\todo{complete table with final result (here it's just false exemple)}
\begin{table}[!h]
	\centering
	\begin{tabular}{cc|ccc}
		\multicolumn{2}{r}{}&\multicolumn{2}{c}{Our result}&\multicolumn{1}{c}{Bentley et. al. 2004}\\
			N&$\mu$ & $\alpha$ (sd) & $p$-value (sd) &$\alpha$ (sd)\\\hline
		250	&0.004&1.53 (0.03)&0.8 (0.05)&1.54 (0.02)\\
			&0.016&1.57 (0.02)&0.6 (0.05)&1.57 (0.01)\\
			&0.064&1.66 (0.01)&0.0 (0.00)&1.67 (0.01)\\\hline
		500	&0.004&1.50 (0.02)&0.6 (0.05)&1.53 (0.03)\\
			&0.016&1.55 (0.03)&0.3 (0.05)&1.61 (0.04)\\
			&0.064&1.78 (0.08)&0.0 (0.00)&1.81 (0.10)\\
	\end{tabular}
	\caption{Mean $\alpha$ \& SD are calculate on 100 runs for our results, and 5 runs for Bentley et. al. 2004.}
	\label{tab:mualpha}
\end{table}

@@

	\end{center}
	\caption{Frequency distribution, where each points represent the mean for 100 runs, for: a) the neutral and the trading models.  b) the neutral model and the trading model without the trading innovation process. c) the trade model and the trade model without the trading innovation process.}
	\label{fig:2setDi}
\end{figure}

On figure~\ref{fig:2setDi}.a) it appears that the implementation of trade mechanisms leads to a distribution of prices departing from the neutral hypothesis. In more details, the frequency distribution has a plateau of common prices (a number of prices share similar and high frequency), which shows that, when trade is taken into account, the most common variants are more diverse.

@@

The successful implementation of a trading model within our framework shows that trading models can be viewed as a particular implementation of the cultural evolutionary framework. This aspect which has not been studied before, opens a new line of thinking on the weight of trade mechanisms within cultural change. Notably, we have shown that the implementation of trade mechanisms leads to a distribution of prices departing from the neutral hypothesis, which is the reference in the study of cultural changes. Within the trading model, the frequency distribution has a plateau of common prices (a number of prices share similar and high frequency), which shows that, when trade is taken into account, the most common variants are more diverse. Interestingly, we have not find references pointing at the ability of trade mechanisms to keep a relatively large diversity in the frequency distribution. We suspect that this is mainly because it is not common to compare results of trading models to other cultural evolution models. It would therefore be interesting to compare the frequency distributions obtained by our model and the ones observed in current or past economies. On the side of cultural transmission, the results obtained can also be compared to the ones obtained when prestige biased cultural transmission is used. The idea behind this last comparison, is that trade could be interpreted as a particular prestige biased cultural transmission and therefore be fully integrated within the cultural evolutionary frameworks.

>>>>

We first analyse the result obtained in the ``neutral model" with one good. The figure~\ref{fig:allMutation} presents the results obtained for two population sizes $N$	 (250 and 500 agents) and $\mu$ varying through three values (0.004,0.016,0.064). The figure is a log-log plot of the average (across all the runs) of the distribution of variants obtained through all experiments. The y-axis shows the proportions of the variants of the prices used during the simulation, the x-axis shows how many variant achieves such proportions.

\begin{figure}[!h]
	\centering
	\begin{tabular}{ c c}
		 $N=250$ & $N=500$ \\
		\includegraphics[width=6cm]{img/allmuRandMaxN250.pdf}&
		\includegraphics[width=6cm]{img/allmuRandMaxN500.pdf}
	\end{tabular}
	\caption{Distribution of proportions depending on the $\mu$ parameter with 250 agents (left) and 500 agents (right). Each plot is the mean obtenaid for 100 runs.\label{fig:allMutation}}
\end{figure}

@@

\todo{complete table with final result (here it's just false exemple)}
\begin{table}[!h]
	\centering
	\begin{tabular}{cc|ccc}
		\multicolumn{2}{r}{}&\multicolumn{2}{c}{Our result}&\multicolumn{1}{c}{Bentley et. al. 2004}\\
			N&$\mu$ & $\alpha$ (sd) & $p$-value (sd - pr) &$\alpha$ (sd)\\\hline
		250	&0.004&1.53 (0.03)&0.58 (0.24 - .01)&1.54 (0.02)\\
			&0.016&1.57 (0.02)&0.35 (0.26 - .05)&1.57 (0.01)\\
			&0.064&1.66 (0.01)&0.0 (0.00 - 1)&1.67 (0.01)\\\hline
		500	&0.004&1.50 (0.02)&0.6 (0.05)&1.53 (0.03)\\
			&0.016&1.55 (0.03)&0.3 (0.05)&1.61 (0.04)\\
			&0.064&1.78 (0.08)&0.0 (0.00)&1.81 (0.10)\\
	\end{tabular}
	\caption{Mean $\alpha$ \& sd are calculate on 100 runs for our results, and 5 runs for Bentley et. al. 2004, pr is the percent of run for which the $p$-value<0.05 (we reject the hypothese that the distribution follow a power law).}
	\label{tab:mualpha}
\end{table}

@@

	\end{center}
	\caption{Proportions distribution, where each points represent the mean for 100 runs, for: a) the neutral and the trading models.  b) the neutral model and the trading model without the trading innovation process. c) the trade model and the trade model without the trading innovation process.}
	\label{fig:2setDi}
\end{figure}

On figure~\ref{fig:2setDi}.a) it appears that the implementation of trade mechanisms leads to a distribution of prices departing from the neutral hypothesis. In more details, the proportion distribution has a plateau of common prices (a number of prices share similar and high proportion), which shows that, when trade is taken into account, the most common variants are more diverse.

@@

The successful implementation of a trading model within our framework shows that trading models can be viewed as a particular implementation of the cultural evolutionary framework. This aspect which has not been studied before, opens a new line of thinking on the weight of trade mechanisms within cultural change. Notably, we have shown that the implementation of trade mechanisms leads to a distribution of prices departing from the neutral hypothesis, which is the reference in the study of cultural changes. Within the trading model, the frequency distribution has a plateau of common prices (a number of prices share similar and high proportion), which shows that, when trade is taken into account, the most common variants are more diverse. Interestingly, we have not find references pointing at the ability of trade mechanisms to keep a relatively large diversity in the frequency distribution. We suspect that this is mainly because it is not common to compare results of trading models to other cultural evolution models. It would therefore be interesting to compare the frequency distributions obtained by our model and the ones observed in current or past economies. On the side of cultural transmission, the results obtained can also be compared to the ones obtained when prestige biased cultural transmission is used. The idea behind this last comparison, is that trade could be interpreted as a particular prestige biased cultural transmission and therefore be fully integrated within the cultural evolutionary frameworks.
